\chapter{Introduction}\label{C:intro}

\section{Problem Statement}
\emph{Service-oriented computing} (SOC) is a novel computing paradigm that employs services as fundamental elements to achieve the agile development of cost-efficient and integrable enterprise applications in heterogeneous environments \cite{papazoglou2003service, papazoglou2006p}. SOC aims to be platform neural and language agnostic to enable integrable and seamless communication among those existing or newly built independent services. \emph{Service Oriented Architecture} (SOA)  could abstractly implement service-oriented paradigm of computing. This accomplishment has been contributing to the reuse of software components, from the concept of functions to units and from units to services during the ongoing development of SOA \cite{booth2004web, overdick2007resource}. One of the most typical implementation of SOA is \emph{web service}, which is designated as ``modular, self-describing, self-contained applications that are available on the Internet" \cite{curbera2001web}. Several standards play a significant role in registering, enquiring and grounding web services across the web, such as UDDI \cite{curbera2002unraveling}, WSDL \cite{lausen2007semantic} and SOAP \cite{fensel2011semantic}. Since an atomic web can not always satisfy users' requirements, \emph{web service composition} aims to loosely couple a set of web services to provide a value-added composite service that accommodates complex requirements of service users. 

The first notable challenges for web service composition are ensuring interoperability of services \cite{fensel2011semantic}. \emph{Interoperability} of web services presents challenges in syntactic and semantic dimensions. On the one hand, the syntactic dimension is covered by the XML-based technologies \cite{yu2008deploying}, such as $WSDL$ and $SOAP$. In this dimension, most services are composed together merely based on the matching of input-output parameters. On the other hand, the semantic dimension enables a better collaboration through ontology-based semantics \cite{o2005review}, in which many standards have been established, such as OWL-S \cite{martin2004owl}, Web Service Modeling Ontology (WSMO) \cite{lausen2005w3c}, SAWSDL \cite{kopecky2007sawsdl}, and Semantic Web Services Ontology (SWSO) \cite{petrie2016web}. This dimension brings around some other underlying functionality of services (i.e., precondition and postcondition) that could effect the execution of web services and their composition. The interoperability challenge gives birth to \emph{Semantic web services composition}, as a different technique from traditional service composition methods (i.e., only syntactic dimension is presented in web services). The interoperability of services is presented as a semantic matchmaking relationship between provided information (i.e.， one service output) and required information (i.e., another service input), which do not often perfectly match with each other based on their semantic descriptions. Therefore, the quality of matchmaking becomes an aspect of the quality concern for users, which also raise researchers' interests. 


The second notable challenges for web service composition are achieving Quality of Service (QoS) optimization \cite{fensel2011semantic}. QoS represents non-functional attributes of service composition (e.g, cost, time, reliability and availability). Often, a global search method is employed to minimize the cost and maximize the reliability of composite services. This challenge gives birth to \emph{QoS-aware service composition} that aims to find composition solutions with optimized QoS. Furthermore, QoS-aware service composition problem is facilitated by \emph{Service Level Agreement (SLA)} \cite{sahai2002automated}, i.e., binding constraints on QoS. This results in the \emph{SLA-aware web service composition}, e.g., constraints on cost, execution time, availability and reliability are separately specified with both lower and upper bounds. In addition, when QoS preference is provided by users, optimization problem often returns a single near-optimal solution, and it is considered as a \emph{single-objective optimization} problem. Otherwise, a set of composition solutions are returned when QoS preference is not clearly by users. This problem is considered as a \emph{multi-objective optimization problem}.  

Apart from the two notable challenges discussed above, the service composition problem is dynamic in the real world, rather than \emph{static}. For example, QoS of services are fluctuating over time. These services chosen at the planning stage may not be available during the runtime, it may due to imprecise interface description \cite{ishikawa2011bridging} or hardware failures \cite{guinard2009discovery}. New services can also be registered in the service repository from time to time. Existing techniques for \emph{static web service composition} cannot cope with such a changing environment. Therefore, \emph{dynamic web service composition} becomes a very important research field with growing research interest and practical value. Particularly, some mechanisms are required to automatically detect the changes or recover from the faults \cite{chan2009fault}. Additionally, in the context of semantic web service composition, dynamic web service composition must also handle the changes in the ontology.


Different service composition approaches \cite{da2016genetic,da2016particle,gupta2015optimization,lecue2009optimizing,ma2015hybrid,qi2010combining,rodriguez2010composition,yu2013adaptive,wang2014automated} have been proposed to cope with those composition challenges discussed above and they can be classified into two main categories: \emph{semi-automated web service composition} and \emph{fully automated web service composition}. The first category of composition approach requires human beings to manually create abstract service workflows. It is generally assumed that a pre-defined abstract workflow is given a priori. The optimization problem in this approach turns to selecting concrete services with the best possible quality for each abstract service in the given workflow. Due to a tremendous growth in industries and enterprise applications, the number of web services has increased dramatically at an unprecedented pace. The process of designing abstract workflows  manually is fraught with difficulties in effectively and efficiently solving composition problems \cite{lecue2009optimizing}, such as QoS-aware service composition problem. Therefore, full automation of composition process is introduced in web service composition for less human intervention, less time consumption, and high productivity \cite{rao2004survey}. The differences in fully automated approach is that an abstract workflow is not provided, but instead created automatically with the service selection process. 


Generating composition plans automatically in discovering and selecting suitable web services is an NP-hard problem \cite{moghaddam2014service}, which means the near-optimized composition solution is not likely to be found in a reasonable computation time. On the one hand, traditional approaches, such as Integer Linear Programming (ILP) and Dynamic Programming. They often suffer scalability problems when the number of services increases and/or the size of ontology increases. On the other hand, to effectively and efficiently solve this problem, \emph{Artificial Intelligence (AI) planning-based approaches}, \emph{Evolutionary Computation (EC) techniques} and hybrid techniques have been introduced. AI planning is utilized to solve the automated web service composition problems as a plan making process, from initial states to a set of actions to desired goal states-composite web services, where services are considered as actions triggered by one state (i.e., inputs) and resulted in another state (i.e., outputs). In the second approach, heuristics have been employed to generate near-optimal solutions using a variety of EC techniques, e.g., Genetic Algorithms (GA), Genetic Programming (GP) and Particle Swarm Optimization (PSO). EC-based techniques have been effectively developed to solve the \emph{QoS-aware web service composition} problem. Many representations have been carefully investigated in QoS-aware web service composition problems, since they could significantly affect the performance of fully automated service composition approaches. In the third approach, a hybridization of AI planning-based approaches and EC-based approaches \cite{da2016genetic,ma2015hybrid} have been proposed to simultaneously ensure the correctness of service compositions and optimize the quality of composition solutions (e.g., QoS). These methods are considered to be more effective in finding more near-optimal solutions. As summarized here, a few challenges previously addressed, but not solved for fully automated service compositions lies in jointly optimizing QoS and semantic matchmaking quality, dealing with multi-objective semantic web service composition for simultaneously optimizing different quality aspects, handling environment changes (i.e., changes in QoS, ontology and service repository) in dynamic service composition, and composing semantic web services considering precondition and effects while optimizing QoS and semantic matchmaking quality. Those challenges are addressed by existing works, but with some key limitations discussed in Section \ref{C:motivation}. 

The overall goal of this thesis is to propose effect and efficient approaches to comprehensive-quality aware automated web service composition. This comprehensive quality aims to jointly optimize semantic matchmaking quality and QoS. Meanwhile, this new approach also tackles several service composition problems, such as multi-objective optimization, dynamic web service composition and semantic web service composition based on preconditions and postconditions.

\section{Motivations}\label{C:motivation}
The motivations of this proposed research lies in five key aspects that simultaneously account for. 
\begin{enumerate*}
 \item \emph{Various techniques of combining AI planning and EC for automatic service composition}.
 \item \emph{Simultaneous handling quality of semantic matchmaking and quality of service}.
 \item \emph{Muti-objective optimization}.
 \item \emph{Dynamic semantic service composition}.
 \item \emph{Service composition based on preconditions and effects}.
\end{enumerate*}
Herein these key aspects are explicitly discussed below. 
\subsubsection{Various Techniques of Hybridization}

Various techniques have been utilized to solve service composition problems, such as AI planning, local searching and EC-based techniques \cite{feng2013dynamic,parejo2008qos,qi2010combining,wang2014automated}. AI planning is a prominent technique for handling web service composition problems while always ensuring the correctness of the solution (i.e., all the inputs of involved services are satisfied) \cite{wang2014automated}.  Local search is a exhaustive search technique for solving optimization problems. In local search, solutions keep moving to the neighbor solutions, driven by some local maximization criterion until near-optimal solution found \cite{parejo2008qos}. However, this technique has the shortage of being trapped in local optimal. On the other hand, EC-based techniques are outstanding at solving combinatorial optimization problems and are less prone to premature convergence in complex search spaces. To take the benefits from various techniques,  various techniques of hybridization allows escaping local optima easily and improving the rate of convergence rate \cite{renders1996hybrid}.

Various techniques have been employed to handle service composition problems in the literature \cite{feng2013dynamic,markou2015non,parejo2008qos,peer2005web,qi2010combining,wang2014automated}. Many researchers investigated AI planning techniques for service composition problems using classical planning algorithms (e.g., GraphPlan \cite{blum1997fast}), where inputs, outputs, preconditions and effects are well defined along with the actions (i.e., services) \cite{markou2015non,peer2005web}. On the one hand, AI planning ensures both the correctness of functionality and satisfactory of constrains, but it is always considered to be less efficient and less scalable, incapable of solving complicated optimization problems \cite{parejo2008qos}. On the other hand, some researchers combined AI planning algorithms and local algorithms search to handle optimization problems, e.g., a combination of Graphplan and Dijkstra’s algorithm is proposed by \cite{feng2013dynamic} to find a correct solution with optimized QoS. EC techniques have been utilized to handle service composition problems, due to their promising performance in handling combinatorial optimization problems in a huge searching space. A few researchers also combined both local search and EC-based techniques for efficiently finding composition solution with optimized QoS \cite{da2016memetic,parejo2008qos}. From the techniques discussed above, they are problem-specific for either optimizing QoS or number of services. In this thesis, more complicated and realistic service composition problems are addressed. To cope with this composition problems featured in all the motivations discussed below (i.e., comprehensive quality-aware service composition, multi-objective optimization, dynamic service composition, and composing semantic web services based on preconditions and postconditions ), EC-based hybrid techniques, optionally combing with AI planning or local search, are motivated to be proposed in our problems. For example, one heuristic in an EC-based technique is initially utilized for a purpose of global search. Meanwhile, another heuristic is additionally designed for a purpose of a local search, which determining neighborhood of composition solutions. This hybrid method may lead to more effective performance.

\subsubsection{Comprehensive Quality of Semantic Web Service Composition}\label{SC:hybridisation}
Web service compositions are optimized by the well known non-functional attributes (i.e., QoS). In the domain of semantic web services, often, the provided information (i.e., outputs) does not perfectly match the required information (i.e., inputs) according to their semantic descriptions \cite{lecue2008optimizing}. Then, the quality of these matches (i.e., quality of semantic matchmaking) become one part of the goal for achieving service compositions with better semantic matchmaking quality \cite{lecue2009optimizing}. Therefore, a comprehensive quality is proposed for simultaneously considering both QoS and semantic matchmaking quality as a novel combinatorial optimization problem. One motivated example from a practical perspective is explained here: many different service compositions can meet a user request but differ significantly in terms of QoS and semantic matchmaking quality. For example, in the classical travel planning context, some component service must be employed to obtain a travel map. Suppose that two services can be considered for this purpose. One service $S$ can provide a street map at a price of 6.72. The other service $S'$ can provide a tourist map at a price of 16.87. Because in our context a tourist map is more desirable than a street map, $S'$ clearly enjoys better semantic matchmaking quality than $S$ but will have negative impact on the QoS of the service composition (i.e., the price is much higher). One can easily imagine that similar challenges frequently occur when looking for service compositions. Hence, a good balance between QoS and semantic matchmaking quality is called for.

Existing works on service composition focus mainly on addressing only one quality aspect discussed above. For the semantic matchmaking quality, it is mainly addressed in works that focus on the discovery of atomic services, i.e., one-to-one matching of user requirements to a single service. Some works \cite{bansal2016generalized,boustil2014semantic,mier2015integrated} on semantic service composition often minimize the number of services or the size of a graph representation for a web service composition. These approaches do not even guarantee an optimized overall QoS. On the other hand, huge efforts have been devoted to studying QoS-aware web service composition \cite{da2015graphevol,da2016particle,gupta2015optimization,ma2015hybrid,qi2010combining,yu2013adaptive}. Some of these works do consider different semantic matchmaking types (e.g., Exact and Plugin matches \cite{paolucci2002semantic}) when they compose services, but do not recognize the importance of semantic matchmaking quality. It is not sufficient to only consider one quality aspect for optimizing service composition. However, only very few works address both semantic matchmaking quality and QoS for web service composition problems. To the best of our knowledge, \cite{fanjiang2014semantic,lecue2009optimizing,pop2009immune} reported about first attempts that consider both aspects together, but only semi-automated approaches have been studied in their works. For these reasons, there is a need to device a comprehensive quality for jointly optimizing the two quality aspects. Apart from that, new fully automated approaches need to be proposed to handle this optimization problem and explicitly support the comprehensive quality. 


\subsubsection{Multi-Objective Composition Optimization}
Existing approaches for handling web service composition problems fall into two groups, depending on different goals of optimization, either a single objective or multiple objectives. In single-objective service compositions, one composition solution is always returned by a composition task, where the preferences of each quality component within the single objective (e.g., a weighted sum of different quality criteria) is known by users. However, users do not always have clear preferences for many quality criteria in advance, and trade-offs presented by two or more conflicting criteria are also taken into optimal decisions. Therefore, multi-objective optimization is a natural requirement to provide a set of trade-off solutions that concern about the conflicting and independent quality criteria. For example, premium users do not care cost as much as price-sensitive users do, so premium users may prefer a composition solution with lowest execution time with no limit on a budget, rather than the one with a relatively lower execution time without exceeding a budget. Therefore, multi-objective web service composition approaches are motivated to be proposed for producing a set of solutions, since the preferences of different quality is not clear and hard to determine in advance, and single-objective optimization using weighted sum method cannot reach solutions in the non-convex regions \cite{kim2006adaptive}.  

Existing research on the automated web service composition mainly concentrates on single objective problems for QoS-aware web service compositions. That is, only one solution is promoted by a unified QoS ranking score to the users. However, to our best knowledge, existing works on multi-objective service composition \cite{liu2005dynamic,wada2012e3,yao2009qos,yin2014hybrid} are only approached by semi-automated methods to handle the conflicting QoS attributes independently. Simultaneously constructing workflows and selecting proper services for optimizing multi-objectives is a very challenge work to complete. On the other hand, some constraints on SLA have also been taken into account in some approaches discussed above. These constraints raise the complexity of absolute Preto priority relation \cite{garey1979guide}. Apart from SLA constraints, users often have vague thoughts about what solutions are preferred. For example, a maximally acceptable trade-off for service execution cost and reliability. These partial preferences should be used for searching the most preferred region. Many preference articulation techniques for multi-objective optimization have been studied \cite{branke2016using,branke2005integrating,branke2001guidance,cheng2015reference,giagkiozis2014pareto}, these approaches can be classified into three groups based on the articulation of preferences \cite{van2000multiobjective}. The first group is priori approaches, where the articulation is done before optimization process; the second group is interactive approaches, where the articulation is done during optimization process; the third group is posteriori approaches, where the articulation is done after optimization process. As argued in \cite{giagkiozis2014pareto}, the third approaches are capable of generating more desired solutions in the preferred region by utilizing biased crowding distance \cite{branke2005integrating} or reference vectors \cite{cheng2015reference}. However, all these approaches have not been used for solving service composition problem considering user preferences on a comprehensive quality. In particular, more practical and effective user preference models are required along with modifications of posteriori methods for effective cope with for service composition problems. From above discussion, these is a lack of fully automated approaches to multi-objective web service composition problems for QoS-aware web service composition abiding by constrains on SLA. Meanwhile, the insufficiency of handling only non-functional attributes (i.e., QoS) has given rise to adding semantic matchmaking quality into simultaneous consideration for this multi-objective problem. Further, user preference articulation techniques have not been studied in service composition problem so far.

\subsubsection{Dynamic Semantic Web Service Composition}
In a dynamic environment, QoS of atomic services in a service repository is fluctuating over time. Static service composition solution is no longer enough, and remedy actions must be taken if the original composition solution changes in QoS or is no longer executable due to any service involved goes offline. Apart from that, newly registered services could also alter the composition plan as it might significantly contribute to the overall QoS or quality of semantic matchmaking. Therefore, dynamic web service composition is proposed to effectively and efficiently update composition solutions when their performance and executability cannot be guaranteed any further \cite{li2014fault}. 


Most approaches work on effective and efficient methods for service re-selection for each component services in a service composition solution, which do not allow the changes of service composition structure. Apart from that, the cost of initial planning is ignored and separated from the adaption of dynamic environment. On the one hand, some techniques \cite{andrews2003business,baresi2011self,koning2009vxbpel} endeavored to update outdated or incorrect compositions, and they allow for dynamic adaptation of the solutions based on implementation of variability constructs at the language level. For example, a composition language extending a typical WS-BPEL \cite{andrews2003business} is proposed for supporting the dynamic adaption using ECA (Event Condition Action) rules, which is utilized for guiding the operations for self-reconfiguration. This approach is difficult to manage, and error-prone. Based on and by extending the previous approaches, variability model based approaches \cite{alferez2014dynamic} have been proposed to support the adaption of service composition solutions. Meanwhile, some other works \cite{mohanty2010web,salas2006ws,wagner2016robust,yin2010qos} utilize decision tree learning, reinforcement learning and Rtree query techniques for dynamic service composition problem. Those non-EC based approaches focus on developing efficient methods for service re-selection, which is utilized for replacing services once negatively contributed changes are detected. On the other hand, EC-based techniques can overcome two limitations. That is, the changes of composition structure is allowed and avoiding re-planning cost through re-using known plans (composite services) obtained prior. First, a proper amount of individuals stored could be used for retrieving an alternative composition solution in the case of failure. This is a save of computation cost for re-planning. Second, the stored individuals could be further evolved resulting in changes of component services or composition structure. However, existing EC-based approaches to web service composition are only studied in static scenarios, rather than dynamic ones. Although some works \cite{feng2013dynamic,liu2005dynamic} claimed that their approaches fit the dynamic problems due to the natural features of their adopted algorithm (i.e., a continuous optimization process), there is no definition or discussion about dynamic service composition problems in their papers. Therefore, there is a lack of research in the area of EC-based dynamic service composition. Given above discussion, it is very advisable to study the effectiveness of EC approaches in a dynamic composition context.


\subsubsection{Automated Web Service Composition Based on Preconditions and Effects}

Since conditions and effects do not apply to most of web services, we achieve valid service composition merely on inputs and outputs previously. However, some web services include conditions and effects, these conditional constraints also determine the executability of services, which lead to more complex service composition constructs (i.e., choice and loop) to be considered when services are composed together. For example, in a scenario of an online book shopping system \cite{wang2014automated}, services are composed to provide an operation for book shopping.  Users expect a purchasing outcome (e.g., receipt) returned if books and customer details (e.g., title, author, customer id) are given. In this case, users may have specific constraints. If the customer has enough money to pay for the book in full amount, then they would like to do so. Otherwise, the customer would like to pay by installments. Therefore, the constraints on their current account balance needs to be handled during the execution of the service composition.

Most of the existing approaches to automated web service composition are approached through services represented by only inputs and outputs. However, a underlying functional knowledge base of services (i.e., in the form of preconditions and effects) is not covered \cite{paliwal2012semantics}. On the one hand, a few approaches \cite{bansal2016generalized,DBLP:journals/soca/BoustilMS14} have been explored to achieve compositions that consider precondition and effects using AI planning, since AI planning ensures both the correctness of functionality and satisfaction on constrains. Meanwhile, exhaustive methods are utilized with AI planning for tackling optimization problems. These methods suffer hugely in terms of efficiency, scalability, and computation cost. On the other hand, EC techniques (i.e., heuristic methods) are considered to be more flexible and more efficient. Given the benefits of both AI planning and EC-based techniques, they are motivated to be collectively explored for automated web service composition based on preconditions and effects.
 
\section{Research Goals}
The overall goal of this thesis is to \emph{develop new and effective EC-based hybrid approaches for comprehensive quality-aware automated semantic web service composition}. More specifically, the research focus will be on: (1) developing EC-based approaches that could explicitly support a proposed comprehensive quality model that jointly optimizes QoS and semantic matchmaking quality, (2) developing multi-objective approaches for optimizing multiple quality criteria involved in the comprehensive quality model, (3) developing EC-based composition techniques to handle various changes in composition environment, and  (4) developing EC-based approaches for semantic web service composition, particularly, considering precondition and effects. This research aims to develop EC-based approaches, optionally combined with some local search, for effectively and efficiently handling the research focus above. The research goal described above can be achieved by completing the following set of objectives:


\begin{enumerate}
  \item \label{Obj:1} \textbf{Develop EC-based approaches to comprehensive quality-aware automated web service composition that simultaneously optimizes both QoS and semantic matchmaking quality}. Particularly, we extend existing works on QoS-aware service composition by considering jointly optimizing both quality aspects, which will be covered collectively through a comprehensive quality model. Meanwhile, representations of the composition solutions are the key aspect of the approaches, since they must maintain all the required information for the evaluation. Therefore, we will investigate the following sub-objectives to handle this objective.
  \begin{enumerate}
    \item \emph{To propose a comprehensive quality model that addresses QoS and semantic matchmaking quality simultaneously to reach a desirable compromise.} We aim to establish a quality model with a simple calculation and good performance. This model is utilized for evaluating QoS and quality of semantic matchmaking of service composition solutions. Meanwhile, to perform a feasible evaluation, it must support most of existing benchmark datasets, e.g., Web Service Challenge 2009 (WSC09)\cite{kona2009wsc} and OWLS-TC \cite{kuster2008opossum}.
    
    \item \emph{To propose direct solution representations for comprehensive quality-aware web service composition.} Graph-based and tree-based representations are widely used for directly representing service composition solutions. Graph-based representations are capable of presenting all the semantic matchmaking relationships as edges, but hardly supporting some composition constructs (e.g., loop and choice). Tree-based representations could be more ideal for practical use, since they can present all composition constructs as inner nodes of trees. However, they could hardly maintain all the edge-related relationships supported by graphs. To take advantage of the graph-based and tree-based representations, we aim to propose a tree-like representation. In particular, any isomorphic copy in the traditional tree-based representations is removed, and insert an edge to the root of the copy. Meanwhile, particular genetic operators will be developed without breaking the functionality of the inserted edge.
    
     \item \emph{To propose indirect solution representations for comprehensive quality-aware web service composition.} The \emph{indirect representations} do not present the final service composition solutions directly, they must be decoded to produce executable service composition solutions. Previous studies have shown their good performance in searching optimal solution for QoS-aware web service composition \cite{da2016memetic,da2016particle}. However, the decoding process could increase the computation time. Apart from that, the indirect representation potentially increases the searching space as discussed in \cite{da2016particle}. To overcome these disadvantages, it is advisable to propose more effective indirect representations.
    
    \item \emph{To propose EC-based hybrid methods to effectively and efficiently handle the problem for comprehensive quality-aware automated web service composition.} The reasons of utilizing hybrid techniques are briefly discussed in Subsection \ref{C:motivation}, and they aim to efficiently and effectively handle service composition problems. Herein, memetic approaches are developed for supporting both the proposed indirect and indirect representations. In particular, new and effective genetic operators are to be developed for EC-based algorithms, and local search techniques will also be developed to avoid being trapped by the local optimal. For the local search, properly designed structures will be studied for neighboring solutions. Apart from that, these approaches must explicitly support the comprehensive quality model. Further more, AI planning techniques will be incorporated into the newly memetic approaches to expedite the convergence speed.
  \end{enumerate}
 
 \item \label{Obj:2} \textbf{Develop multi-objective approaches to fully automated comprehensive quality-aware semantic service composition}. In practice, many quality criteria proposed in our comprehensive quality model are often conflicting in natural, such as an amount of money paid for service execution and a possibility of service availability. Existing works \cite{chen2014partial,xiang2014qos,yin2014hybrid,liu2005dynamic,yu2013efficient,zhang2010qos} mainly concentrated on semi-automated QoS-aware web service composition. Therefore, research must be carried out not only for a better understanding of inherent trade-offs among different objectives (e.g., quality of semantic matchmaking and QoS are naturally considered as two conflicting objectives), but also for developing fully automated approaches for overcoming the limitations of semi-automated approaches. It is approached by utilizing cutting-edge multi-objective optimization algorithms (e.g, NSGA-II \cite{deb2002fast}, SPEA2 \cite{zitzler2001spea2} and MOEA/D \cite{zhang2007moea}). These algorithms are needed for finding a Pareto front of evolved solutions. Meanwhile, different representations may not perform equally well, so a study on improving the effectiveness of different representations along with different Evolutionary Multi-Objective Optimization (EMO) techniques also arouses researchers' interest. Apart from that, SLA consideration needs to be taken into account. It is also necessary to consider customized matchmaking levels to bring the flexibility for meeting different requirements from segmented users (e.g., platinum users, gold users and normal users). Further more, Most EMO algorithms work on achieving evenly distributed non-dominated Pareto solutions. However, users are often interested in certain regions of the Pareto Front, and often have vague ideas about what solutions will be mostly preferred. Therefore, more effective and efficient methods should be conducted for determining the potentially preferred solutions and yielding a smaller set of most preferred solutions. The development of this approach can be divided into the following four sub-objectives:
   \begin{enumerate}
   
    \item \label{Obj:2.1} \emph{To develop a EMO-based approaches for multi-objective fully automated semantic web service composition}. Herein we will propose useful modifications of multi-objective EC algorithms (for example, NSGA-II \cite{deb2002fast}, SPEA2 \cite{zitzler2001spea2} and MOEA/D \cite{zhang2007moea}) simultaneously to improve the effectiveness and efficiency of our service composition approaches. This sub-objective is also established for in-depth investigation of each quality criteria based on our proposed comprehensive quality model in Objective \ref{Obj:1}.  In particular, both quality of semantic matchmaking and QoS must be optimized separately, since they can conflict each other. It would be interesting to examine different tradeoffs among the service composition solutions with respect to the different quality criterion. Apart from that, fully automated approaches are also developed to achieve high practical usefulness without replying on given abstract workflows.
   
    \item \emph{To develop EMO-based hybrid approaches for multi-objective fully automated semantic web service composition}. Once we achieve the sub-objective 2(a), the effectiveness and efficiency are the next focus. The EC-based approaches should be extended by introducing an effective local search mechanism. In particular, we aim to design an effective mechanism to decide how often, how long, which solutions we use the local search for targeting the neighboring solutions. For example, a local search is designed and to target the neighborhood of the best the individual of each current population, and it is performed with a highest possibility with respect to its fitness values. 

    \item \emph{To develop EMO-based approaches for multi-objective fully automated semantic web service composition subject to constraints on SLA and customized matchmaking level}. In real world, sometimes, satisfaction on given SLA constraints is required in addition to optimizing QoS. Therefore, this sub-objective consider some additional constraints on  QoS (i.e., multilevel constraints with lower and upper bound for different individual QoS criterion \cite{yin2014hybrid}). Meanwhile, to satisfy the customized different semantic matchmaking levels (e.g., exact matchmaking level and less strict matchmaking level), these methods are also required to cope with the constraints on the different accepted matchmaking level. We aim to propose a vector that represents QoS and semantic matchmaking constraints, which is modeled as a set of lower and upper bounds for each quality criterion. To effectively handle the constrains, the vector could be utilized in the fitness function for penalizing the individuals that breach the constraints.

    \item \emph{To develop EMO-based approaches for multi-objective fully automated semantic web service composition subject to preference articulation}.  The articulation of user preferences could be achieved either before (i.e., a priori), during (i.e., interactive), or after (i.e., a posteriori) an EMO process. As argued in \cite{giagkiozis2014pareto}, a posteriori preference articulation often results in more and more desired solutions on the preferred region of a Pareto Front. We aim to develop fully automated multi-objective semantic web service composition approach integrating posteriori preference articulation methods. In particular, an effective preference model is created to cope with full or partial preference information according to users' preferences. Meanwhile, this preference model must be explicitly supported by posteriori articulation techniques,  which guild the solutions towards both true Pareto Front and the most preferred region. 

   \end{enumerate}
   
\vspace{0.5cm}
 \item \textbf{Develop EC-based techniques to achieve dynamic semantic web service composition effectively}. Objectives \ref{Obj:1} and \ref{Obj:2} are proposed for solving static service composition problems. In our context, composition environment refers to the registered services in the service repository, non-functional attributes advertised by service providers, and the ontology utilized for describing the resources of web services. On the one hand, after using existing EC-based approaches to generate an initial composition plan from a given composition task, some factors could significantly impact the execution of the plan, such as changes in QoS, ontology and service repository. On the other hand, existing non-EC based approaches \cite{nasridinov2012qos,salas2006ws,wagner2016robust,yin2010qos} work on effective and efficient methods for service re-selection for repairing each service employed. Technically, their approaches do not allow changes in service composition structure. Also, the cost of initializing a composition plan is ignored and separated from their approaches. To overcome the two limitations above, three studies are conducted as three sub-objectives below:

  \begin{enumerate}
 \item \label{Obj:3.1} \emph{To develop EC-based techniques to re-optimize solution candidates in response to changes in QoS and Ontology.} Traditionally, initial population is created with solution candidates that are further evolved for searching optimal solutions. During the evolutionary computation process, most of service candidates are discarded except the best service candidate identified. Those discarded solution candidates may be promising, since some of them could turn to be alternative best due to the changes in services or the ontology. Therefore, instead of discarding the solution candidates, we always create an archive for remaining candidates, which are utilized to re-optimization since these candidates preserve both diversity and elitism. This diversity and elitism motivate us to continue the evolutionary process without wasting the computation time in the stage, where an initial composition plan is produced.

 \item \emph{To develop EC-based hybrid techniques to re-optimize solution candidates for changes in QoS and Ontology.} Once the EC-based techniques to re-optimize solution candidates for changes in QoS and ontology are achieved, it should be further studied in developing more effective approach to handle this problem. We aim to propose an adaptive and hybrid approach to this dynamic problem. On the one hand, we firstly assign a higher priority to a group of services with changes. This priority is utilized in service selection for building up new individual (i.e., service composition solutions), and must be adaptively handled with a proper decreasing rate with respect to each service, when they are actually used in service selection. On the other hand, the hybrid refers to a memetic approach that combines EC-based approach and a local search, where the local search is developed to accelerate the convergency and decrease the number of evaluation times. In this sub-objective, we aim to achieve more effective performance compared to the EC-based approaches in Objective 3(a).
 
 
 \item \emph{To develop EC-based hybrid techniques for handling service failure and new service registration using updated candidates in the population.} Apart from the changes in the QoS and the ontology of services, occasionally, existing services may fail and/or new service may become available. For the case of new service registrations, efficient methods need to be proposed to update the plan with new services in timely manner. When new services are registered in the service repository, we discard a portion of the current population. This discarded population is associated with low fitness values. Then, we replenish population based on updated services repository. For the case of service failure, efficient approaches need to be proposed to either mutate on an un-invokable component service of an individual, or a pre-defined parent structure of the component service, or effectively re-generate whole solutions using invokable services in the service repository. 
 \end{enumerate}
   
 \item \textbf{Develop EC-based approaches for semantic web service compositions based on preconditions and effects (Optional)}. We plan to extend the service composition approaches (i.e., satisfaction on inputs and outputs) to include preconditions and effects. These conditional constraints also necessitates the study of various of composition constructs for automated semantic web service composition, e.g., loop and choice. Therefore, three sub-objectives have been proposed as our targets as follow.
 \begin{enumerate}
 
  \item \emph{To develop EC-based techniques for semantic web service composition based on preconditions and effects}. In this objective, inputs and outputs are not everything for the execution of web services, our composition problem is extended by considering not only inputs and outputs, but also preconditions and effects. In particular, we need to establish a general matchmaking mechanism for satisfaction on preconditions and effects, since there is no systematic studying for that in existing works. We aim to model inputs and outputs as variable names and further referenced in preconditions and effects, which can be distinguished as different instances from the knowledge base. In particular, preconditions are assigned to the description of requirements of inputs using logic formula. The effects are restricted to the description of constraints on returned outputs, relationships between inputs and outputs, and changes caused by the service in the knowledge bases. Based on the mechanism, two simple composition constructs (i.e., sequence and parallel) are automatically constructed for considering preconditions and effects. Meanwhile, we aim to develop EC-based approaches to effectively handle this problem. In particular, we will propose new representations to copy with preconditions and effects.

  \item \emph{To develop EC-based hybrid techniques for semantic web service composition based on preconditions and effects}. Once the EC-based techniques to semantic web service composition based on preconditions and effects are proposed. More effective techniques shall be developed. In particular, we aim to create a memetic approach that utilize a EC-based approach and a local search for avoiding trapped by local optimal. This memetic approach is expect to benefit in an accelerated convergency rate and a decreasing the number of evaluation times.
    
   \item \emph{To develop EC-based techniques for semantic web service composition based on preconditions and effects for supporting loops and choice}. We initially extend the matchmaking mechanism of satisfaction on preconditions and effects to include loops and choice composition constructs. To extensively cope with these two constructs, new and effective representations must be studied since the old representation only consider sequence and parallel constructs. Apart from that, EC-based approaches will be developed to effectively solve this problem.

 
 \end{enumerate}
 
\end{enumerate}

\section{Published Papers}

During the initial stage of this research, the preliminary work was carried out on establishing the comprehensive quality model.  Afterwards, some studies on the direct and indirect representations are completed for one part of Objective \ref{Obj:1}, but the earlier works focus on static web service composition using single-objective optimization technique. The following are the publications made from the preliminary studies:

\begin{itemize}
 \item WANG, C., MA, H., CHEN, A., AND HARTMANN, S. ''Comprehensive Quality-Aware Automated Semantic Web Service Composition``. \textit{AI 2017: Advances in Artificial Intelligence: 30th Australasian Joint Conference}. 2017, pp. 195-207.
 \item WANG, C., Ma, H., CHEN, A., AND HARTMANN, S.: ''GP-Based Approach to Comprehensive quality-aware automated semantic web service composition``. In: SEAL2017: International Conference on Simulated Evolution and Learning(To appear)
\end{itemize}


\section{Organization of Proposal}The remainder of the proposal is organized as follows: Chapter \ref{C:review} provides a fundamental definition of the web service composition problem and performs a literature review covering a range of works in this field; Chapter \ref{C:preliminary} presents a formal definition for our service composition problem, and discusses the preliminary work explores direct and indirect representations for comprehensive quality-aware semantic web service composition using a hybridization of AI planning techniques and EC-based techniques; Chapter \ref{C:plan} presents a plan detailing this project's intended contributions, a project timeline, and a thesis outline.
